A generalized definition of multidimensional item response theory parameters

### 2024-04-09 output/paper_draft/paper_M2PLM_parameters.Rmd
Implements changes approved in Google Docs

Our aim is to find in the first place
the point of maximum slope of $`r IRF_ABBR`$ in $`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.
However, Reckase's Equation 3 implicitly assumes orthogonality.
In order to contemplate the general case,
we need to introduce a few linear algebra results before proceeding further.

@@

Let $`r TEST_SPACE`$ be the vector space made up by
the set of $`r DISCR_VECTOR`$ vectors,
with the sum and product by a scalar operation as in $`r REAL_N_SPACE`$.
Consider also an orthonormal basis $`r TEST_SPACE_STD_BASIS`$
in $`r TEST_SPACE`$, such that $`r DISCR_STD_COORDS_ALT_EQ`$.
Since the response probability to the `r MODEL_ACRONYM`,
given by \@ref(eq:M2PL-formulation), must be invariant with respect to
the change of basis expressed in \@ref(eq:change-basis),
we must define the coordinates $`r DISCR_STD_COORDS`$ of the parameter vector
$`r DISCR_VECTOR`$ in the orthonormal basis $`r TEST_SPACE_STD_BASIS`$,
such that $`r INVARIANCE_EQ`$.
If we assume that $`r DISCR_VECTOR`$ exists in vector space $`r TEST_SPACE`$,
with coordinates $`r DISCR_COORDS`$ in basis $`r TEST_SPACE_BASIS`$,
then we must define $`r DISCR_COORDS`$ such that

@@

We want @reckase_difficulty_1985's definition of the multidimensional parameters
in rectangular coordinates to be a particular case of our more general
definition, so we can conveniently assume that the inner product in
$`r TEST_SPACE_STD_BASIS`$ is the dot product,
as we did with $`r LS_ORTH_BASIS`$.
The norm of $`r DISCR_VECTOR`$ thus fulfills

@@

We shall refer to this new space as the *test space* [@zhang_theoretical_1999].
Although the inverse relationship between their respective inner product
Gram matrices induces a bijective relationship between the two sets of spaces,
it is worth highlighting that test space $`r TEST_SPACE`$
and latent space $`r LATENT_SPACE`$ are different entities. Only in the
orthonormal case the two spaces share a common inner product
(i.e., the dot product), and thus can be mistaken.

>>>>

Following @reckase_difficulty_1985’s procedure,
we must find in the first place
the point of maximum slope of $`r IRF_ABBR`$ in $`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.
However, Reckase's Equation 3 implicitly assumes orthogonality.
In order to contemplate the general case,
we need to introduce a few linear algebra results before proceeding further.

@@

Let $`r TEST_SPACE`$ be the vector space made up by
the set of $`r DISCR_VECTOR`$ vectors,
with the sum and product by a scalar operation as in $`r REAL_N_SPACE`$.
We want @reckase_difficulty_1985's definition of
the multidimensional parameters in rectangular coordinates to be
a particular case of our more general definition;
we can thus consider an orthonormal basis $`r TEST_SPACE_STD_BASIS`$
in $`r TEST_SPACE`$ such that $`r DISCR_STD_COORDS_ALT_EQ`$
corresponds in the model to $`r TRAIT_ORTH_COORDS`$.
Since the response probability to the `r MODEL_ACRONYM`,
given by \@ref(eq:M2PL-formulation), must be invariant with respect to
the change of basis expressed in \@ref(eq:change-basis),
Therefore, we find a basis $`r TEST_SPACE_BASIS`$ in $`r TEST_SPACE`$
defined by

@@

with $`r TRANSFORM_MATRIX_TRANSP_INV`$ a change of basis matrix.
Note that, as $`r TEST_SPACE_STD_BASIS`$ is orthonormal,
$`r TRANSFORM_MATRIX_TRANSP_INV`$ determines $`r TEST_SPACE_BASIS`$.

To make @reckase_difficulty_1985's definition a particular case of ours
we can conveniently assume that the inner product in
$`r TEST_SPACE_STD_BASIS`$ is the dot product,
as we did with $`r LS_ORTH_BASIS`$.
The norm of $`r DISCR_VECTOR`$ thus fulfills

@@

We shall refer to this new space $`r TEST_SPACE`$ as
the *test space* [@zhang_theoretical_1999].
Although the inverse relationship between their respective inner product
Gram matrices induces a bijective relationship between the two sets of spaces,
it is worth highlighting that test space $`r TEST_SPACE`$
and latent space $`r LATENT_SPACE`$ are different entities. Only in the
orthonormal case the two spaces share a common inner product
(i.e., the dot product), and thus can be mistaken.
